---
title: "Summary of Mussel Outcomes"
author: "Bradley Saul"
date: "July 9, 2015"
output: pdf_document
---
```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(elktoe)
#river$date_ <- mdy(river$date)
elktoe <- subset(mussels_wide, species == 'A. raveneliana')
elktoe <- elktoe[order(elktoe$river, decreasing = TRUE), ]
```


## Baseline comparisons by River and Species
- Quick boxplot summaries to see if randomization balance the weight and volumes of mussels by river per species.
- Buoyant weights were not available for L. fasciola. Dry weights not available for A. raveneliana.

## Baseline: Buoyant weight

```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot('buoyant_weight_g_0')
```

## Baseline: Buoyant weight: by site
```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot_site('buoyant_weight_g_0', 'A. raveneliana')
```

## Baseline: Dry weight

```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot('dry_weight_g_0')
```

## Baseline: Dry weight: by site

```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot_site('dry_weight_g_0', 'L. fasciola')
```

## Baseline: Volume

```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot('volume_0')
```

## Baseline: Volume: A. raveneliana: by site

```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot_site('volume_0', 'A. raveneliana')
```

## Baseline: Volume: L. fasciola: by site

```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot_site('volume_0', 'L. fasciola')
```

## Baseline: comments

On quick glance, each species appears fairly balanced in each river. The number of mussels are small, so the differences in boxplots aren't surprising. 

## Pre-Post Changes
- All changes are paired differences between first measurements (4/6/13) and last measurements (either 8/19/13, 9/19/13, or 5/6/14). 
- Missing values are noted in the tables and missingness may be due to death of a mussel or other reasons. 
- Fewer values are missing for volume because L/W/H measurements could be taken on dead organisms.

## Change: Buoyant Weight
```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot('buoyant_weight_g_d')
```

## Change: Buoyant Weight
```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'asis'}
cat("Species | River | Mean | Min | Median | Max | Missing | Dead | %Dead ", 
    "--- | --- | --- | --- | --- | --- | --- | --- | ---", sep="\n")
cat(apply(summarize_mussel('buoyant_weight_g_d', c('species', 'river')), 
          1, function(X) paste(X, collapse=" | ")), sep = "\n")
```


```{r, echo=FALSE, warning=FALSE}
ar <- mussels_wide[mussels_wide$species == 'A. raveneliana', ]
wilcox.test(ar$buoyant_weight_g_d ~ ar$river)
```


## Change: Buoyant Weight: by site
```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot_site('buoyant_weight_g_d', 'A. raveneliana')
```

## Change: Dry Weight
```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot('dry_weight_g_d')
```

## Change: Dry Weight
```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'asis'}
cat("Species | River | Mean | Min | Median | Max | Missing | Dead | %Dead ", 
    "--- | --- | --- | --- | --- | --- | --- | --- | ---", sep="\n")
cat(apply(summarize_mussel('dry_weight_g_d', c('species', 'river')), 
          1, function(X) paste(X, collapse=" | ")), sep = "\n")
```

```{r, echo=FALSE, warning=FALSE}
lf <- mussels_wide[mussels_wide$species == 'L. fasciola', ]
wilcox.test(lf$dry_weight_g_d ~ lf$river)
```

## Change: Dry Weight: by site
```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot_site('dry_weight_g_d', 'A. raveneliana')
```

## Change: Volume
```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot('volume_d')
```

## Change: Volume
```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'asis'}
cat("Species | River | Mean | Min | Median | Max | Missing | Dead | %Dead ", 
    "--- | --- | --- | --- | --- | --- | --- | --- | ---", sep="\n")
cat(apply(summarize_mussel('volume_d', c('species', 'river')), 
          1, function(X) paste(X, collapse=" | ")), sep = "\n")
```

## Change: Volume
Statistical tests for changes in volume (median):

A. ravenenliana
```{r, echo=FALSE, warning=FALSE}
wilcox.test(ar$volume_d ~ ar$river)
```

L. fasciola
```{r, echo=FALSE, warning=FALSE}
wilcox.test(lf$volume_d ~ lf$river)
```

## Change: Volume: A. raveneliana: by site

```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot_site('volume_d', 'A. raveneliana')
```

## Change: Volume: L. fasciola: by site

```{r, echo=FALSE, warning=FALSE, message=FALSE}
mussel_boxplot_site('volume_d', 'L. fasciola')
```

## Change: Comments

- Outliers should be investigated. Are these true outliers or were measurements incorrectly recorded?
- L. fasciola gained dry weight in both rivers. 
- A. raveneliana lost weight in both rivers - there may be a difference between the two rivers in loss of buoyant weight. One has to keep in mind that dead mussels did not contribute to weight measurements.

## A. raveneliana Weight

```{r, echo=FALSE,  warning = FALSE, message = FALSE}
mussel_plot('A. raveneliana', 'buoyant_weight_g')
```
    
## A. raveneliana volume
```{r, echo=FALSE, warning = FALSE, message = FALSE}
mussel_plot('A. raveneliana', 'volume')
```
     
## L. fasciola weight

```{r, echo=FALSE, warning = FALSE, message = FALSE}
mussel_plot('L. fasciola', 'dry_weight_g')
```

## L. fasciola volume
```{r, echo=FALSE, warning = FALSE, message = FALSE}
mussel_plot('L. fasciola', 'volume')
```

## A. raveneliana: % mortaliy:
```{r, echo=FALSE, warning = FALSE, message = FALSE}
temp <- ddply(mussels_wide, .(species, river, site), summarize,
              mortality = mean(dead, na.rm= T))
temp$site_order <- with(temp, substr(site, 6, 6))
ggplot(temp, aes(x = site_order, y = mortality)) + geom_bar(stat = 'identity') +
  facet_grid(river ~ species) + 
  theme_classic()

```

## A. raveneliana: lost 10% of BW or died:
```{r, echo=FALSE, warning = FALSE, message = FALSE}
temp <- ddply(elktoe, .(river), summarize,
              mean_lost_dead = mean(lost_10_or_dead, na.rm= T))
mussel_barchart(temp, 'river', 'mean_lost_dead', '% lost 10% of body weight or died')
```

## A. raveneliana: lost 10% of BW or died: by site
```{r, echo=FALSE, warning = FALSE, message = FALSE}
temp <- ddply(elktoe, .(river, site), summarize,
              mean_lost_dead = mean(lost_10_or_dead, na.rm= T))
mussel_barchart(temp, 'site', 'mean_lost_dead', '% lost 10% of body weight or died')
```


## A. raveneliana: change in body weight by corbicula density: by site
```{r, echo=FALSE, warning = FALSE, message = FALSE}
cc <- coef(lm(buoyant_weight_g_d ~ corb_density_0, data = elktoe))

ggplot(elktoe, aes(x = corb_density_0, y=buoyant_weight_g_d, color = river)) +
    geom_point() + geom_hline(yintercept = 0) +
    scale_fill_brewer(guide = 'none', palette = 'Set2', c) +
    geom_abline(intercept = cc[1], slope = cc[2], color = 'blue') + 
    xlab('Corbicula Density') + ylab('Change in buoyant weight (g)') +
    theme_classic() +
    theme(legend.key = element_blank(), legend.title = element_blank())
```

## A. raveneliana: proportion change in body weight by corbicula density: by site
```{r, echo=FALSE, warning = FALSE, message = FALSE}
cc <- coef(lm(p_buoyant_weight_d ~ corb_density_0, data = elktoe))

ggplot(elktoe, aes(x = corb_density_0, y=p_buoyant_weight_d, color = river)) +
    geom_point() + geom_hline(yintercept = 1) +
    scale_fill_brewer(guide = 'none', palette = 'Set2', c) +
    geom_abline(intercept = cc[1], slope = cc[2], color = 'blue') + 
    xlab('Corbicula Density') + ylab('Proportion change in buoyant weight (g)') +
    theme_classic() +
    theme(legend.key = element_blank(), legend.title = element_blank())
```


## A. raveneliana: % died by corbicula density
```{r, echo=FALSE, warning = FALSE, message = FALSE}
temp <- ddply(elktoe, .(river, corb_density_0, site), summarize,
              mean_dead = mean(dead, na.rm= T))
ggplot(temp, aes(x = corb_density_0, y=mean_dead, color = river, 
                 label = site)) +
    geom_point() + geom_text(hjust=.5, vjust=-1, size = 3.5) + 
    scale_color_brewer(guide = 'none', palette = 'Set2') +
    xlab('Corbicula Density') + ylab('% Died') +
    theme_classic() +
    scale_y_continuous(limits = c(0, 1)) +
    theme(legend.key = element_blank(), legend.title = element_blank())
```

## A. raveneliana: % dies or lost any weight by corbicula density
```{r, echo=FALSE, warning = FALSE, message = FALSE}
temp <- ddply(elktoe, .(river, corb_density_0, site), summarize,
              mean_lost_dead = mean(lost_wt_or_dead, na.rm= T))
ggplot(temp, aes(x = corb_density_0, y=mean_lost_dead, color = river, 
                 label = site)) +
    geom_point() + geom_text(hjust=.5, vjust=-1, size = 3.5) + 
    scale_color_brewer(guide = 'none', palette = 'Set2') +
    xlab('Corbicula Density') + ylab('% Died or lost any weight') +
    theme_classic() +
    scale_y_continuous(limits = c(0, 1)) +
    theme(legend.key = element_blank(), legend.title = element_blank())
```

## A. raveneliana: % dies or lost more than 10% of BW by corbicula density
```{r, echo=FALSE, warning = FALSE, message = FALSE}
temp <- ddply(elktoe, .(river, corb_density_0, site), summarize,
              mean_lost_dead = mean(lost_10_or_dead, na.rm= T))
ggplot(temp, aes(x = corb_density_0, y=mean_lost_dead, color = river, 
                 label = site)) +
    geom_point() + geom_text(hjust=.5, vjust=-1, size = 3.5) + 
    scale_color_brewer(guide = 'none', palette = 'Set2') +
    xlab('Corbicula Density') + ylab('% Died or lost > 10% of BW') +
    theme_classic() +
    scale_y_continuous(limits = c(0, 1)) +
    theme(legend.key = element_blank(), legend.title = element_blank())
```
